---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
raw_data <- read.csv(file = 'activity.csv')
head(raw_data)
str(raw_data)
```


## What is mean total number of steps taken per day?
Total number of steps taken per day:
```{r}
steps_each_day <- aggregate(x = raw_data$steps, by = list(raw_data$date), FUN = sum)
names(steps_each_day) <- c('date', 'steps')
head(steps_each_day)
str(steps_each_day)
```

Histogram of the total number of steps taken each day:
```{r}
hist(steps_each_day$steps, main = 'Histogram of the total number of steps taken by day', xlab = 'Steps')
```

Mean of the total number of steps taken per day:
```{r}
mean(steps_each_day$steps, na.rm = TRUE)
```

Median of the total number of steps taken per day:
```{r}
median(steps_each_day$steps, na.rm = TRUE)
```


## What is the average daily activity pattern?
```{r}
by_interval <- aggregate(raw_data$steps, list(raw_data$interval), mean, na.rm = TRUE)
names(by_interval) <- c('date', 'steps')
plot(by_interval$date, 
     by_interval$steps, 
     type = 'l', 
     main = 'Average number of steps taken, averaged across all days', 
     xlab = '5-minute interval', ylab = 'Avg. number of steps taken')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
by_interval[by_interval$steps == max(by_interval$steps), ]
```


## Imputing missing values
Total number of missing values in the dataset:
```{r}
sum(is.na(raw_data$steps))
```


## Are there differences in activity patterns between weekdays and weekends?
