---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r message=FALSE}
library(dplyr)
library(ggplot2)
```


## Loading and preprocessing the data
```{r}
raw_data <- read.csv(file = 'activity.csv')
head(raw_data)
tail(raw_data)
str(raw_data)
```


## What is mean total number of steps taken per day?
Total number of steps taken per day:
```{r}
steps_each_day <- aggregate(x = raw_data$steps, by = list(raw_data$date), FUN = sum)
names(steps_each_day) <- c('date', 'steps')
head(steps_each_day)
str(steps_each_day)
```

Histogram of the total number of steps taken each day:
```{r}
hist(steps_each_day$steps, main = 'Histogram of the total number of steps taken by day', xlab = 'Steps')
```

Mean of the total number of steps taken per day:
```{r}
mean(steps_each_day$steps, na.rm = TRUE)
```

Median of the total number of steps taken per day:
```{r}
median(steps_each_day$steps, na.rm = TRUE)
```


## What is the average daily activity pattern?
```{r}
by_interval <- aggregate(raw_data$steps, list(raw_data$interval), mean, na.rm = TRUE)
names(by_interval) <- c('interval', 'steps')
plot(by_interval$interval, 
     by_interval$steps, 
     type = 'l', 
     main = 'Average number of steps taken, averaged across all days', 
     xlab = '5-minute interval', ylab = 'Avg. number of steps taken')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
by_interval[by_interval$steps == max(by_interval$steps), ]
```


## Imputing missing values
Total number of missing values in the dataset:
```{r}
sum(is.na(raw_data$steps))
```

Filling in missing values using the mean of the 5-minute interval:
```{r}
nas <- raw_data[is.na(raw_data$steps), ]
nas <- tbl_df(nas)
by_interval <- tbl_df(by_interval)
nas_filled <- left_join(nas, by_interval, by = c('interval' = 'interval'))
nas_filled <- select(nas_filled, date, interval, steps.y)
names(nas_filled) <- c('date', 'interval', 'steps')

raw_no_nas <- raw_data
raw_no_nas[is.na(raw_no_nas$steps), ]$steps <- nas_filled$steps
```

Total number of steps taken per day:
```{r}
steps_each_day_no_nas <- aggregate(x = raw_no_nas$steps, by = list(raw_no_nas$date), FUN = sum)
names(steps_each_day_no_nas) <- c('date', 'steps')
```

Comparing histograms of raw data versus NA's filled:
```{r}
par(mfrow = c(1, 2))
hist(steps_each_day$steps, main = 'Raw data', xlab = 'Steps')
hist(steps_each_day_no_nas$steps, main = 'Na\'s filled', xlab = 'Steps')
par(mfrow = c(1, 1))
```

Mean of the total number of steps taken per day (NA's filled):
```{r}
mean(steps_each_day_no_nas$steps)
```

Median of the total number of steps taken per day (NA's filled):
```{r}
median(steps_each_day_no_nas$steps)
```

Filling in missing values seems to make no change in the mean and 
median summaries.

Watching the histograms, the more frequent steps count seems to have now 
a higher frequency.

## Are there differences in activity patterns between weekdays and weekends?

Calculating the mean for each 5-minute interval on weekdays and weekends:
```{r}
by_interval_no_nas <- tbl_df(raw_no_nas)
by_interval_no_nas <- mutate(by_interval_no_nas, 
                             date = as.Date(date),
                             type = ifelse(
                                     weekdays(date) == 'Domingo' | 
                                             weekdays(date) == 'Sábado', 
                                     'weekend', 'weekday')) %>% 
        group_by(type, interval) %>% 
        summarise(steps = mean(steps))
ggplot(by_interval_no_nas, aes(interval, steps)) + geom_line() + facet_wrap(~type, nrow = 2)
```
